Simple Trial Report

### 2022-04-11 inst/jss/listdown-jss.tex
Typo.

\Abstract{
Reproducible document standards, like \proglang{R} Markdown, facilitate
the programmatic creation of documents whose content is itself
programmatically generated. While programmatic content alone may not be
sufficient for a rendered document since it does not include prose
(content generated by an author to provide context, a narrative, etc.)
programmatic generation can provide substantial efficiencies for
structuring and constructing documents. This paper explores the
programmatic generation of reproducible documents by distinguishing
components that can be created by computational means from those
requiring human-generation, providing guidelines for the generation of
these documents, and identifying a use case in clinical trial reporting.
These concepts and use case are illustrated through the \pkg{listdown}
package for the \proglang{R} programming environment, which is is
currently available on the Comprehensive R Archive Network (CRAN).
}

@@

\begin{itemize}
\item{\bf as\_ld\_yml() }{- turn a computational component list into YAML with class information}
\item{\bf ld\_cc\_dendro() }{- create a dendrogram from a list of computational components}
\item{\bf ld\_chunk\_opts() }{- apply chunk options to a presentation object}
\item{\bf ld\_ioslides\_header() }{- create an ioslides presentation header}
\item{\bf ld\_make\_chunks() }{- write a listdown object to a string}
\item{\bf ld\_rmarkdown\_header() }{- create an R Markdown header}
\item{\bf ld\_workflowr\_header() }{- create a worflowr header}
\item{\bf ld\_write\_file() }{- write to and R Markdown file}
\item{\bf listdown() }{- create a listdown object to create an \proglang{R} Markdown document}
\item{\bf print.listdown()}{- print the listdown options for \proglang{R} Markdown document creation}
\end{itemize}

The rest of this paper is structured as follows. The next section goes
over basic usage and commentary. This section is meant to convey the
approach used by the package and shows how to describe an output
document using \pkg{listdown}, create a document, and change how the
presentation of computational components can be specialized using
\pkg{listdown} decorators. With the user accustomed to the package's
basic usage, section 3 describes the design of the package. Section 4
goes over advanced usage of the package including adding initialization
code to and outputted document as well as how to control chunk-level
options. Section 5 provides a simplified case study of how the package
is currently being used in clinical trial reporting. Section 6 concludes
the paper with a few final remarks on the general types of applications
where \pkg{listdown} has been shown effective.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library(ggplot2)
R> 
R> library(listdown)
R> 
R> data(anscombe)
R> 
R> computational_components <- list(
+   Linear = ggplot(anscombe, aes(x = x1, y = y1)) + geom_point(),
+   `Non Linear` = ggplot(anscombe, aes(x = x2, y = y2)) + geom_point(),
+   `Outlier Vertical`= ggplot(anscombe, aes(x = x3, y = y3)) + 
+     geom_point(),
+   `Outlier Horizontal` =  ggplot(anscombe, aes(x = x4, y = y4)) + 
+     geom_point())
R> 
R> ld_cc_dendro(computational_components)
\end{CodeInput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> ld <- listdown(load_cc_expr = readRDS("comp-comp.rds"), 
+                package = "ggplot2",
+                echo = FALSE)
R> 
R> ld_make_chunks(ld)[1:7]
\end{CodeInput}
\begin{CodeOutput}
[1] ""                                     
[2] "```{r echo = FALSE}"                  
[3] "library(ggplot2)"                     
[4] ""                                     
[5] "cc_list <- readRDS(\"comp-comp.rds\")"
[6] "```"                                  
[7] ""                                     
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> saveRDS(comp_comp2, "comp-comp2.rds")
R> ld <- listdown(load_cc_expr = readRDS("comp-comp2.rds"),
+                package = c("ggplot2", "DT", "purrr"),
+                decorator = list(ggplot = identity,
+                                 data.frame = datatable_no_search),
+                setup_expr = knitr::opts_chunk$set(echo = FALSE),
+                init_expr = {
+                  datatable_no_search <- partial(datatable,
+                                                 options = list(dom = 't'))
+                  })
R> 
R> ld_make_chunks(ld)[2:14]
\end{CodeInput}
\begin{CodeOutput}
 [1] "```{r setup, include = FALSE}"                                         
 [2] "knitr::opts_chunk$set(echo = FALSE)"                                   
 [3] "```"                                                                   
 [4] ""                                                                      
 [5] "```{r}"                                                                
 [6] "library(ggplot2)"                                                      
 [7] "library(DT)"                                                           
 [8] "library(purrr)"                                                        
 [9] ""                                                                      
[10] "datatable_no_search <- partial(datatable, options = list(dom = \"t\"))"
[11] ""                                                                      
[12] "cc_list <- readRDS(\"comp-comp2.rds\")"                                
[13] "```"                                                                   
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library(gtsummary)
R> library(dplyr)
R> library(survival)
R> library(survminer)
R> library(rmarkdown)
R> 
R> make_surv_cc <- function(trial, treat, surv_cond_chars) {
+   table_1 <- trial %>%
+     tbl_summary(by = all_of(treat)) %>%
+     gtsummary::as_flex_table()
+ 
+   scs <- lapply(c("1", surv_cond_chars),
+                 function(sc) {
+                   sprintf("Surv(ttdeath, death) ~ %s + %s", treat, sc) %>%
+                     as.formula() %>%
+                     surv_fit(trial) %>%
+                     ggsurvplot()
+                 })
+   names(scs) <- c("Overall", tools::toTitleCase(surv_cond_chars))
+   list(`Table 1` = table_1, `Survival Plots` = scs)
+ }
R> 
R> surv_cc <- make_surv_cc(trial, treat = "trt",
+                         surv_cond_chars = c("stage", "grade"))
R> 
R> ld_cc_dendro(surv_cc)
\end{CodeInput}
\begin{CodeOutput}

@@

\bibliography{../references}

>>>>

\Abstract{
Reproducible document standards, like \proglang{R} Markdown, facilitate
the programmatic creation of documents whose content is itself
programmatically generated. While programmatic content alone may not be
sufficient for a rendered document since it does not include prose
(content generated by an author to provide context, a narrative, etc.)
programmatic generation can provide substantial efficiencies for
structuring and constructing documents. This paper explores the
programmatic generation of reproducible documents by distinguishing
components that can be created by computational means from those
requiring human-generation, providing guidelines for the generation of
these documents, and identifying a use case in clinical trial reporting.
These concepts and use case are illustrated through the \pkg{listdown}
package for the \proglang{R} programming environment, which is is
currently available on the Comprehensive R Archive Network.
}

@@

\begin{itemize}
\item{\bf as\_ld\_yml() }{- turn a computational component list into YAML with class information.}
\item{\bf ld\_cc\_dendro() }{- create a dendrogram from a list of computational components.}
\item{\bf ld\_chunk\_opts() }{- apply chunk options to a presentation object.}
\item{\bf ld\_ioslides\_header() }{- create an ioslides presentation header.}
\item{\bf ld\_make\_chunks() }{- write a listdown object to a string.}
\item{\bf ld\_rmarkdown\_header() }{- create an R Markdown header.}
\item{\bf ld\_workflowr\_header() }{- create a worflowr header.}
\item{\bf ld\_write\_file() }{- write to and R Markdown file.}
\item{\bf listdown() }{- create a listdown object to create an \proglang{R} Markdown document.}
\item{\bf print.listdown()}{- print the listdown options for \proglang{R} Markdown document creation.}
\end{itemize}

The rest of this paper is structured as follows. The next section goes
over basic usage and commentary. It is meant to convey the
approach used by the package and shows how to describe an output
document using \pkg{listdown}, create a document, and change how the
presentation of computational components can be specialized using
\pkg{listdown} decorators. With the user accustomed to the package's
basic usage, Section 3 describes the design of the package. Section 4
goes over advanced usage of the package including adding initialization
code to and outputted document as well as how to control chunk-level
options. Section 5 provides a simplified case study of how the package
is currently being used in clinical trial reporting. Section 6 concludes
the paper with a few final remarks on the general types of applications
where \pkg{listdown} has been shown effective.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library("ggplot2")
R> 
R> library("listdown")
R> 
R> data(anscombe)
R> 
R> computational_components <- list(
+   Linear = ggplot(anscombe, aes(x = x1, y = y1)) + geom_point(),
+   `Non Linear` = ggplot(anscombe, aes(x = x2, y = y2)) + geom_point(),
+   `Outlier Vertical`= ggplot(anscombe, aes(x = x3, y = y3)) + 
+     geom_point(),
+   `Outlier Horizontal` =  ggplot(anscombe, aes(x = x4, y = y4)) + 
+     geom_point())
R> 
R> ld_cc_dendro(computational_components)
\end{CodeInput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> ld <- listdown(load_cc_expr = readRDS("comp-comp.rds"), 
+                package = "ggplot2",
+                echo = FALSE)
R> 
R> ld_make_chunks(ld)[1:7]
\end{CodeInput}
\begin{CodeOutput}
[1] ""                                     
[2] "```{r echo = FALSE}"                  
[3] "library(\"ggplot2\")"                     
[4] ""                                     
[5] "cc_list <- readRDS(\"comp-comp.rds\")"
[6] "```"                                  
[7] ""                                     
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> saveRDS(comp_comp2, "comp-comp2.rds")
R> ld <- listdown(load_cc_expr = readRDS("comp-comp2.rds"),
+                package = c("ggplot2", "DT", "purrr"),
+                decorator = list(ggplot = identity,
+                                 data.frame = datatable_no_search),
+                setup_expr = knitr::opts_chunk$set(echo = FALSE),
+                init_expr = {
+                  datatable_no_search <- partial(datatable,
+                                                 options = list(dom = 't'))
+                  })
R> 
R> ld_make_chunks(ld)[2:14]
\end{CodeInput}
\begin{CodeOutput}
 [1] "```{r setup, include = FALSE}"                                         
 [2] "knitr::opts_chunk$set(echo = FALSE)"                                   
 [3] "```"                                                                   
 [4] ""                                                                      
 [5] "```{r}"                                                                
 [6] "library(\"ggplot2\")"
 [7] "library(\"DT\")" 
 [8] "library(\"purrr\")"
 [9] ""                                                                      
[10] "datatable_no_search <- partial(datatable, options = list(dom = \"t\"))"
[11] ""                                                                      
[12] "cc_list <- readRDS(\"comp-comp2.rds\")"                                
[13] "```"                                                                   
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library("gtsummary")
R> library("dplyr")
R> library("survival")
R> library("survminer")
R> library("rmarkdown")
R> 
R> make_surv_cc <- function(trial, treat, surv_cond_chars) {
+   table_1 <- trial %>%
+     tbl_summary(by = all_of(treat)) %>%
+     gtsummary::as_flex_table()
+ 
+   scs <- lapply(c("1", surv_cond_chars),
+                 function(sc) {
+                   sprintf("Surv(ttdeath, death) ~ %s + %s", treat, sc) %>%
+                     as.formula() %>%
+                     surv_fit(trial) %>%
+                     ggsurvplot()
+                 })
+   names(scs) <- c("Overall", tools::toTitleCase(surv_cond_chars))
+   list(`Table 1` = table_1, `Survival Plots` = scs)
+ }
R> 
R> surv_cc <- make_surv_cc(trial, treat = "trt",
+                         surv_cond_chars = c("stage", "grade"))
R> 
R> ld_cc_dendro(surv_cc)
\end{CodeInput}
\begin{CodeOutput}

@@

\bibliography{references}

### 2022-04-11 inst/jss/supplemental-material/listdown-jss.rmd
Typo.

---
documentclass: jss
author:
  - name: Michael Kane
    affiliation: Yale University
    address:
    - 60 College Street
    - New Haven, CT 06510, USA
    email: michael.kane@yale.edu
  - name: Xun (Tony) Jiang
    affiliation: Amgen Inc.
    address:
    - One Amgen Center Drive
    - Thousand Oaks, CA 91320-1799, USA
    email: xunj@amgen.com
  - name: Simon Urbanek
    affiliation: The University of Aukland
    address:
    - 38 Princes Street
    - Auckland Central, Auckland 1010, NZ
    email: s.urbanek@auckland.ac.nz
title:
  formatted: "On the Programmatic Generation of Reproducible Documents"
  # If you use tex in the formatted title, also supply version without
  plain:     "On the Programmatic Generation of Reproducible Documents"
  # For running headers, if needed
  short:     "On the Programmatic Generation of Reproducible Documents"
abstract: >
  Reproducible document standards, like \proglang{R} Markdown, facilitate the programmatic creation of documents whose content is itself programmatically generated. While programmatic content alone may not be sufficient for a rendered document since it does not include prose (content generated by an author to provide context, a narrative, etc.) programmatic generation can provide substantial efficiencies for structuring and constructing documents. This paper explores the programmatic generation of reproducible documents by distinguishing components that can be created by computational means from those requiring human-generation, providing guidelines for the generation of these documents, and identifying a use case in clinical trial reporting. These concepts and use case are illustrated through the \pkg{listdown} package for the \proglang{R} programming environment, which is is currently available on the Comprehensive R Archive Network (CRAN).
keywords:
  # at least one keyword must be supplied
  formatted: [reproducibility, reproducible documents, data presentation]
  plain:     [reproducibility, reproducible documents, data presentation]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

@@

```{r cc1}
library(ggplot2)

library(listdown)

@@

The \code{listdown()} function facilitates the insertion of setup and initialization code through the \code{setup_expr} and \code{init_expr} arguments. If and argument is provided to the \code{setup_expr}, then the first code chunk of the document will have the specified code inserted. This code chunk is named "setup" and the include parameter is set to \code{FALSE}. When the \code{init_expr} argument is specified,
code is inserted immediately after the libraries are loaded in the R Markdown document. In general, it is suggested that the number of initial expressions be kept small so that the \proglang{R} Markdown document is easy to read. If a large number of functions are required by the target \proglang{R} Markdown document then they can be put into a file and sourced using the initial expression. As an example, suppose we are creating an html document and presenting data using the \code{datatable()} function. However, we do not want to include the search capabilities provided. This can be easily accomplished by creating a new function, \code{datatable_no_search()}, created using the \code{partial()} function \cite{purrr} to partially apply \code{list(dom = 't')} to the \code{options} argument of \code{datatable}.

@@

```{r, eval = TRUE, message=FALSE, warning=FALSE}
library(gtsummary)
library(dplyr)
library(survival)
library(survminer)
library(rmarkdown)

>>>>

---
documentclass: jss
author:
  - name: Michael Kane
    affiliation: Yale University
    address:
    - 60 College Street
    - New Haven, CT 06510, USA
    email: michael.kane@yale.edu
  - name: Xun (Tony) Jiang
    affiliation: Amgen Inc.
    address:
    - One Amgen Center Drive
    - Thousand Oaks, CA 91320-1799, USA
    email: xunj@amgen.com
  - name: Simon Urbanek
    affiliation: The University of Aukland
    address:
    - 38 Princes Street
    - Auckland Central, Auckland 1010, NZ
    email: s.urbanek@auckland.ac.nz
title:
  formatted: "On the Programmatic Generation of Reproducible Documents"
  # If you use tex in the formatted title, also supply version without
  plain:     "On the Programmatic Generation of Reproducible Documents"
  # For running headers, if needed
  short:     "On the Programmatic Generation of Reproducible Documents"
abstract: >
  Reproducible document standards, like \proglang{R} Markdown, facilitate the programmatic creation of documents whose content is itself programmatically generated. While programmatic content alone may not be sufficient for a rendered document since it does not include prose (content generated by an author to provide context, a narrative, etc.) programmatic generation can provide substantial efficiencies for structuring and constructing documents. This paper explores the programmatic generation of reproducible documents by distinguishing components that can be created by computational means from those requiring human-generation, providing guidelines for the generation of these documents, and identifying a use case in clinical trial reporting. These concepts and use case are illustrated through the \pkg{listdown} package for the \proglang{R} programming environment, which is is currently available on the Comprehensive R Archive Network.
keywords:
  # at least one keyword must be supplied
  formatted: [reproducibility, reproducible documents, data presentation]
  plain:     [reproducibility, reproducible documents, data presentation]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

@@

```{r cc1}
library("ggplot2")

library("listdown")

@@

The \code{listdown()} function facilitates the insertion of setup and initialization code through the \code{setup_expr} and \code{init_expr} arguments. If an argument is provided to the \code{setup_expr}, then the first code chunk of the document will have the specified code inserted. This code chunk is named "setup" and the include parameter is set to \code{FALSE}. When the \code{init_expr} argument is specified,
code is inserted immediately after the libraries are loaded in the R Markdown document. In general, it is suggested that the number of initial expressions be kept small so that the \proglang{R} Markdown document is easy to read. If a large number of functions are required by the target \proglang{R} Markdown document then they can be put into a file and sourced using the initial expression. As an example, suppose we are creating an html document and presenting data using the \code{datatable()} function. However, we do not want to include the search capabilities provided. This can be easily accomplished by creating a new function, \code{datatable_no_search()}, created using the \code{partial()} function \cite{purrr} to partially apply \code{list(dom = 't')} to the \code{options} argument of \code{datatable}.

@@

```{r, eval = TRUE, message=FALSE, warning=FALSE}
library("gtsummary")
library("dplyr")
library("survival")
library("survminer")
library("rmarkdown")
